The framing of active school travel in Ontario, Canada, as a health and built environment issue

### 2022-01-28 AST-Framing-Ontario/AST-Framing-Ontario.Rmd
Antonio's edits of latest version of paper

---
title: The framing of active school travel in Ontario, Canada as a health and built environment issue
author:
  - name: Author 1
    email: author1@example.com
    affiliation: Some Department
    footnote: 1
  - name: Author 2
    email: author2@example.com
    affiliation: Some Department
  - name: Author 3
    email: author3@example.com
    affiliation: Another University
    footnote: 2
  - name: Author 4
    email: author4@example.com
    affiliation: Some Institute
    footnote: 2
  - name: Author 5
    email: author5@example.com
    affiliation: Some University
  - name: Author 6
    email: author6@example.com
    affiliation: Some Department
address:
  - code: Some Department
    address: Department, Street, City, Province, Postal Code
  - code: Another University
    address: Department, Street, City, Province, Postal Code
  - code: Some Institute
    address: Street, City, Province, Postal Code
  - code: Some University
    address: Department, Street, City, Province, Postal Code
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.

@@

journal: "Social Science & Medicine"
date: "`r Sys.Date()`"
bibliography: bibliography.bib
linenumbers: false
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

@@

Rates of walking and bicycling to school, commonly known as active school travel (AST), have been declining in Canada and the United States for decades [@rothmanDeclineActiveSchool2018]. However, the appetite for AST may be high in many Canadian communities. For instance, 40% of children in a study conducted in Toronto, Ontario who are driven to school would like to travel by bicycle instead [@laroucheRatherBikeSchool2016]. The study also reported that less than 3% of respondents actually cycled, despite the vast majority having access to a bicycle and living within a short and bikeable distance to school. Another study based in London, Ontario reported similar findings with respect to children's preference for active travel [@larsenRouteBasedAnalysisCapture2012]. From a public health perspective, this represents a major opportunity to improve health and wellbeing if more children could use active modes to school.

In response to this trend, the Government of Ontario, in Canada, created a fund in 2017 to support communities across the province in developing AST initiatives. Green Communities Canada runs the program, called _Ontario Active School Travel_ [@GreenCommunities2016]. As of December 2021, OAST has awarded over $2 million and provided resources to over 25 projects across Ontario. Many communities implemented school travel planning (STP), along with encouraging activities like walking school buses and developed resources for schools and parents, among other actions [see @GreenCommunitiesProjects]. In this context, STP is a popular "school-specific" intervention led by a facilitator who brings together a committee of stakeholders from diverse sectors including education, planning, transportation, and public health to develop action plans [@buliungSchoolTravelPlanning2011; @mammenSchoolTravelPlanning2014]. The five-step process involves identifying barriers to AST based on the local context and implementing approaches or activities following the 4E's that make AST safer and more convenient: 1) setup of the program; 2) data collection and problem identification; 3) action planning; 4) implementation; 5) monitoring and evaluation [@buliungSchoolTravelPlanning2011; @langUnderstandingModalChoice2011].

@@

In this paper, we use natural language processing to examine how three STP stakeholder groups in Ontario (municipalities, school boards, and transportation consortia) frame the issue of AST. We assembled a corpus of texts from webpages that present information for the general public or parents who are interested in AST. We examine word frequency, bigrams, and concordances in these selected documents, and also identify key topics presented by each stakeholder group. We then compare the findings from these documents to a selection of studies on AST and explore the extent to which there is concordance between the literature on AST and materials shared with the public.

@@

The benefits of AST are valuable information to communicate to the public in order to convey the importance of this issue. The desire to increase AST in Canada is certainly warranted - there is compelling evidence that children who actively commute to school earn physical and mental health benefits. Faulkner et al. [-@faulknerActiveSchoolTransport2009] concluded from their systematic review that children who travel on foot or by bicycle to school generally have higher levels of physical activity than their peers who are driven to school. However, Schoeppe et al. [-@schoeppeAssociationsChildrenActive2015] reported no association between AST and physical activity in their cross-sectional study. This relationship could be dose-dependent, meaning that children would have to travel longer distances to accumulate physical activity. A walking distance of 1000-1600 metres to school has been found to contribute to overall levels of physical activity for boys [@faulknerSchoolTravelChildren2013]. The daily routine of travelling to school can be a good opportunity for children to regularly build physical activity into their schedule [@mitraIndependentMobilityMode2013]. Research has also shown that using active modes to school contributes to improvements in cardiovascular fitness [@borrestadExperiencesRandomisedControlled2012].

More recently, there has been literature exploring the link between transport and children's wellbeing [@waygoodTransportChildrenWellbeing2020], with relevant applications to the study of travel satisfaction [@vandenbergFactorsAffectingParental2020; @westmanChildrenTravelSchool2017]. Being driven to school reduces community interactions for children which may negatively impact their social wellbeing [@waygoodChildrenTravelIncidental2015], whereas walking and bicycling give children opportunities to socialize with friends or siblings [@michailChildrenExperiencesTheir2021]. This is something that children value highly about travel [@zwertsHowChildrenView2010] which indicates that social connections through travel are important for their wellbeing. AST also provides opportunities for children to engage with the natural environment [@fuscoUnderstandingChildrenPerceptions2012; @romeroChildrenExperiencesEnjoyment2015].

@@

Factors that influence AST have been presented and organized using a socio-ecological model (SEM) [@mitraIndependentMobilityMode2013] or systems model [@badlandDevelopmentSystemsModel2016] whereby children's travel behaviour is understood within the context of household, social, neighbourhood, and policy environments. The SEM comes from the field of public health and is a useful framework for understanding complex health behaviours, including travel habits such as walking or bicycling to school, because it identifies multiple determinants that need to be addressed by interventions to facilitate behaviour change. The consensus in the literature is that interventions should target multiple levels in order to increase levels of healthy and physically active modes of travel to school [@mitraIndependentMobilityMode2013]. Rather than describing an extensive list of factors that influence AST and mode choice in Canada [e.g., @mammenUnderstandingDriveEscort2012; @mitraIndependentMobilityMode2013; @rothmanDeclineActiveSchool2018; @wilsonUnderstandingChildParent2018], we discuss a few potentially modifiable determinants that may be targeted for change through STP.

At the individual level, older child age is often associated with AST [@mammenUnderstandingDriveEscort2012; @starkExploringIndependentActive2018; @wilsonUnderstandingChildParent2018]. There is evidence that gender is a determinant of AST, with boys being more likely to travel using active modes than girls, although this is not a strong or consistent finding [@rothmanDeclineActiveSchool2018; @schoeppeAssociationsChildrenActive2015]. Children's mode choice to school is strongly influenced by their parents' travel behaviours and the complexity of their household's travel needs [@buliungLivingJourneySchool2021], as well as the frequency of parental support [@mahDoesParentalSupport2017a], which indicates that shifting parental perceptions and habits is important. Convenience and inclement weather have been cited by parents as barriers to AST [@buliungSchoolTravelPlanning2011]. Parental perceptions of the built or school environment [@demeesterParentalPerceivedNeighborhood2014; @panterAttitudesSocialSupport2010] and their children's skills [@mammenUnderstandingDriveEscort2012; @faulknerWhatQuickestEasiest2010] also influence mode choice to school.

Distance between home and school is most strongly associated with AST [@ikedaAssociationsChildrenActive2018; @mammenUnderstandingDriveEscort2012; @pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018] with less AST reported among children who have to travel farther to school. Many studies have also found that the quality of the built environment along the route to school and around the school site [@ikedaAssociationsChildrenActive2018; @rothmanActiveSchoolTransportation2021] and provision of active travel infrastructure [@chenPromotingActiveStudent2018; @pontEnvironmentalCorrelatesChildren2009] facilitate AST. Canadian youth report that they feel most safe bicycling on streets in their neighbourhood or that have low volumes of traffic [@TACbikeinfra2020]. Finally, concerns about traffic and strangers have been reported by parents who drive their children to school [@mammenUnderstandingDriveEscort2012], which highlights that the volume and speed of cars can be a concern or deterrent for AST.

@@

How AST is framed seems particularly important to shift parental attitudes and perceptions given their reported influence on children's travel mode to school. STP activities heavily focus on education or encouragement [@buliungSchoolTravelPlanning2011; @mammenActiveSchoolTravel2014; @buttazzoniSupportingActiveSchool2018], but parents may not always be receptive to the goals of STP and may be resistant to behaviour change [@buttazzoniSupportingActiveSchool2018]. Parents have been found to express different understandings, language, and perceptions than planners of the built environment determinants that influence school travel [@buliungLivingJourneySchool2021]. This is also true when it comes to other factors like convenience of different modes to school [@langUnderstandingModalChoice2011]. STP stakeholders must pay special attention to parents' understanding of the decline of AST as a problem, which may affect their receptivity to proposed solutions.

The “central organizing idea or story line" of AST, to apply the definition of framing from Gamson and Modigliani [-@gamsonChangingCulture1987], could also affect broader support in the community. Municipal representatives are perceived to be instrumental but the involvement of other stakeholder groups (e.g., busing consortium representatives and local residents) can be lacking [@buttazzoniSupportingActiveSchool2018]. In Ontario, it would be reasonable to say that AST has become a policy issue on the education and public health agendas over this time as evidenced by the financial contribution from the provincial government. The support from a range of municipal representatives [see @buttazzoniSupportingActiveSchool2018; @mammenPuttingSchoolTravel2015] demonstrates that the AST issue is on the political agenda. However, it is unknown to what degree the general public (i.e., local residents) has been exposed to messaging about AST.

@@

kable(data_tbl,
      "latex",
      booktabs = TRUE,
      caption = "\\label{tab:search-results}Search results from the main STP stakeholder groups.") %>%
  kable_paper(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(3, border_left = T) %>%
  kable_styling(latex_options = c("striped", "scale_down"))
```

@@

__To be completed by Dr. Paez__

@@

A multi-step process was conducted to ensure that the analysis captured as much text as possible from both the policy documents (n = 64) and academic papers (n = 233). To begin, the webpages, which were manually downloaded in portable document format (PDF), were trimmed so that pages that only consisted of tables, figures, or references were removed. Many academic papers were in a two-column format, which is not ideal for conversion to `txt`. We adapted a procedure (https://stackoverflow.com/questions/42541849/extract-text-from-two-column-pdf-with-r) to read the two-column PDF documents so that they would be converted correctly. Four academic papers did not join sufficiently and were taken out of the corpus due to the substantial time required to manually correct their inconsistencies.

Next, we converted the trimmed PDF documents into `txt` files so that they could be imported in R for analysis. We then proceeded to a manual cleaning phase where we removed any remaining tables, figures, references, headers/footings, and captions that could not be trimmed. Manual corrections were also required for certain pages in academic papers that remained in two-column format after the conversion process. This typically occurred on pages that had a table or figure which disrupted the text. Finally, we reviewed all of the documents to remove hyphenation by line breaks and to keep hyphenated words together on the same line. Any ligatures (e.g., combinations of characters or letters that were not properly detected during the conversion process) were fixed.

We also manually removed any extraneous material in the academic papers that did not pertain to AST specifically. This included footnotes, references, acknowledgments, and conflict of interest statements in the academic papers. We removed all phone numbers, inserted links to other webpages, personal names, and content not to specific to AST from the policy documents that were retrieved from the websites of school boards, municipalities, and transportation consortia.

@@

We use topic modelling in R to conduct the framing analysis. Topic modelling is a machine learning technique that can identify what language and concepts are being communicated by analyzing text. This method is practical for researchers working with large amounts of text because it replaces the manual coding of topics that would normally take place to analyze or summarize textual data [@jacobiQuantitativeAnalysisLarge2016]. In the data pre-processing phase, we tokenize the text in the documents and create a document-term matrix so that it is in the correct format for analysis. We primarily use the following packages: `tidytext` [@R-tidytext], `topicmodels` [@R-topicmodels], `word2vec` [@R-word2vec], and `wordcloud` [@R-wordcloud] to examine text in the documents that were sourced for this project. These packages have functions for determining the frequency of specific words in each document or relationships (e.g., pairs of adjacent terms called bigrams) and correlations between words. These methods can reveal what language and concepts are being communicated to the general public. Topic modelling is a popular method for analyzing text from social media platforms [@albalawiUsingTopicModeling2020] and news articles [@jacobiQuantitativeAnalysisLarge2016]. We estimate latent Dirichlet allocation (LDA) models to classify both the STP and academic documents according to the topics that are contained within them. This method "treats each document as a mixture of topics, and each topic as a mixture of words" [@silgeTextMining2021]. The model's output is "a set of topics consisting of clusters of words that co-occur in these documents according to certain patterns" [@jacobiQuantitativeAnalysisLarge2016]. Researchers must then interpret the identified topics, as done after other methods of manual coding. We also compare the topics between the policy documents and the academic papers, and use our interpretation to answer the questions outlined above that guide this paper. This methodology is not familiar to many, so we describe in more detail below how we used specific functions and code in R to produce our results.

@@

This paper is an example of open and reproducible research that uses only open software. All data were obtained from publicly available sources and organized in the form of a data package. Following best practices in spatial data science [@brunsdon2020opening], the code and data needed to reproduce or conduct a similar analysis for other regions in North America or elsewhere are available for download.

@@

We analyzed word and document frequency for each set of documents (i.e., corpus). Table \ref{tab:word-table} shows the most frequent terms found in the municipal, transportation consortia, school board, and academic documents. As expected, policy documents and academic papers reference _active_, _travel_, _walking_, _biking_ or _cycling_, and _students_ more than other terms. Each corpus also has _safety_ and _traffic_ as common words which suggests that such concerns are a common reason given for parents driving their children to school. The word _physical_ is present in each corpus, but it's not clear what this refers to (e.g, _physical activity_, _physical health_, or the _physical environment_). Furthermore, documents from STP stakeholder groups discuss _resources_, _information_, and _services_ about school travel. Unlike the academic papers, policy documents include the words _route_ or _routes_ as frequent terms. This could indicate the presence of "safe routes to school" in the policy corpus, as well as the role of STP stakeholder groups in identifying safe routes to school to share with parents or families. In the section below, the context in which these terms are used is explored further.

The academic corpus differs from the policy documents in that _parents_ and _distance_ are the second and third most common terms. In addition, _time_, _factors_, _environment_, and _age_ are also identified in the academic papers. The prevalence of these terms is consistent with an academic focus on exploring the variables that influence mode choice. These words are absent from the list of common words in policy documents. Table \ref{tab:word-table} indicates that the academic corpus discusses a broader range of determinants of AST than the policy documents. The number of references for each term in the academic papers is also significantly higher due to the inclusion of more documents.

@@

However by examining document frequency, we found that some terms are not present in all policy documents. This suggests that although documents pertain to the subject of school travel, some stakeholders across Ontario are not disseminating information about AST. For example, a document may discuss active travel but not to school. Some documents present school travel options by bus but not by active modes. We manually searched the policy corpus and found that 48% of documents mention AST and 16% mention STP. This confirms that many municipalities, school boards, and transportation consortia are not promoting AST through their webpages or indicating their involvement in the STP process. Instead, inclement weather and its impacts on busing is a common topic addressed in school board and transportation consortia documents.

@@

Bigrams refer to a pair of consecutive words. We used the `unnest_tokens` function from the _tidytext_ package to determine the bigrams. Figures \ref{fig:city-visual}, \ref{fig:consortia-visual}, and \ref{fig:school-visual} show the bigrams that occur more than 5 times for each set of policy documents. These figures help to make further sense of the word frequencies reported above, and highlight the main ideas that are presented to the public in each of the policy corpora. The directional arrows indicate the arrangement of the words (e.g., active travel and not travel active) and the colour gradient of the arrows corresponds to the most frequently mentioned pairs (e.g., bigrams with darker arrows are found more often).

Municipalities primarily discuss _physical activity_ (n = 53) and _public health_ (n = 19) in the context of AST. In addition, _travel planning_ (n = 19), _bike lanes_ (n = 16), and _safe routes_ (n = 14) are also identified, conceivably as either proposed solutions or built environment factors that support AST. Key issues related to transport such as _traffic safety_ (n = 10), _air quality_ (n = 9), and _greenhouse gases_ (n = 9) are conveyed to the public through these policy documents. It is not surprising to find this focus given that municipalities in Ontario are concerned about climate change and have increasingly looked to active modes of travel to offset transport-related emissions in urban areas.

Similar word bigrams are found in school board documents: _travel planning_ (n = 33), _safe routes_ (n = 15), _physical activity_ (n = 10), and _public health_ (n = 10) are among the most common bigrams. Both municipalities and school boards in Ontario seem to emphasize what can be or has been done to improve AST (i.e., policy or planning changes), while outlining some of the benefits of AST at the individual- or community-level to potentially encourage behaviour change (i.e., physical activity for children or improved air quality). Unlike other STP stakeholders, school boards also consider _inclement weather_ (n = 24) and _bus cancellations_ (n = 13). This is likely because many students in Ontario travel to school by bus and this information is presented alongside AST options. Finally, transportation consortia documents highlight topics such as _physical activity_ (n = 10), _pedestrian safety_ (n = 8), _crossing guards_ (n = 6), _travel planning_ (n = 6), and _walk zones_ (n = 6). Biking and cycling are notably absent from transportation consortia documents. Overall, the policy documents appear to convey an emphasis on the built environment, rather than household decision-making.

@@

We interpreted the most common bigrams from the policy corpus (see Figure \ref{fig:policy-visual}), which includes all documents from municipalities, transportation consortia, and school boards, as the main ideas that STP stakeholder groups are focusing on and communicating to the public about AST. We then used the `kwic` function from the _quanteda_ package to better understand the context of these key ideas. Table \ref{tab:policy-concordance} presents some examples of the context that were extracted from select policy documents to demonstrate how the most common bigrams are communicated to the public.

>>>>

---
title: The framing of active school travel in Ontario, Canada, as a health and built environment issue
author:
  - name: Author 1
    email: author1@example.com
    affiliation: Some Department
    footnote: 1
  - name: Author 2
    email: author2@example.com
    affiliation: Some Department
  - name: Author 3
    email: author3@example.com
    affiliation: Another University
    footnote: 2
  - name: Author 4
    email: author4@example.com
    affiliation: Some Institute
    footnote: 2
  - name: Author 5
    email: author5@example.com
    affiliation: Some University
  - name: Author 6
    email: author6@example.com
    affiliation: Some Department
address:
  - code: Some Department
    address: Department, Street, City, Province, Postal Code
  - code: Another University
    address: Department, Street, City, Province, Postal Code
  - code: Some Institute
    address: Street, City, Province, Postal Code
  - code: Some University
    address: Department, Street, City, Province, Postal Code
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.

@@

journal: "Social Science & Medicine"
date: "`r Sys.Date()`"
bibliography: [bibliography.bib, packages.bib]
linenumbers: false
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

@@

Rates of walking and bicycling to school, commonly known as active school travel (AST), have been declining in Canada and the United States for decades [@rothmanDeclineActiveSchool2018]. However, the appetite for AST may be high in many Canadian communities. For instance, in a study conducted in Toronto, Ontario, 40% of children who are driven to school would like to travel by bicycle instead [@laroucheRatherBikeSchool2016]. The study also reported that less than 3% of respondents actually cycled, despite the vast majority having access to a bicycle and living within a short and bikeable distance to school. Another study based in London, Ontario, reported similar findings with respect to children's preference for active travel [@larsenRouteBasedAnalysisCapture2012]. From a public health perspective, increasing the use of active modes to school represents a major opportunity to improve children's health and wellbeing.

The Government of Ontario, in Canada, has striven to promote this trend, and created a fund in 2017 to support communities across the province to develop AST initiatives. Green Communities Canada runs the program, called _Ontario Active School Travel_ [@GreenCommunities2016]. As of December 2021, OAST has awarded over CAD 2 million and provided resources to over 25 projects across Ontario. Many communities implemented school travel planning (STP), along with encouraging activities like walking school buses and developed resources for schools and parents, among other actions [see @GreenCommunitiesProjects]. In this context, STP is a popular "school-specific" intervention led by a facilitator who brings together a committee of stakeholders from diverse sectors including education, planning, transportation, and public health to develop action plans [@buliungSchoolTravelPlanning2011; @mammenSchoolTravelPlanning2014]. The five-step process involves identifying barriers to AST based on the local context and implementing approaches or activities following the 4E's that make AST safer and more convenient: 1) setup of the program; 2) data collection and problem identification; 3) action planning; 4) implementation; 5) monitoring and evaluation [@buliungSchoolTravelPlanning2011; @langUnderstandingModalChoice2011].

@@

In this paper, we use natural language processing to examine how three STP stakeholder groups in Ontario (municipalities, school boards, and transportation consortia) frame the issue of AST. We assembled a corpus of texts from public sources that present information for the general public or parents who are interested in AST. We examine word frequency, bigrams, and concordances in these selected documents, and also identify key topics presented by each stakeholder group. We then compare the findings from these documents to a selection of studies on AST and explore the extent to which there is concordance between the literature on AST and materials shared with the public.

@@

There is value in communicating the benefits of AST to the public in order to convey the importance of this issue. The desire to increase AST in Canada is certainly warranted - there is compelling evidence of the physical and mental health benefits that children who actively commute to school attain. Faulkner et al. [-@faulknerActiveSchoolTransport2009] concluded from their systematic review that children who travel on foot or by bicycle to school generally have higher levels of physical activity than their peers who are driven to school. However, Schoeppe et al. [-@schoeppeAssociationsChildrenActive2015] reported no association between AST and physical activity in their cross-sectional study. This relationship could be dose-dependent, meaning that children would have to travel longer distances to accumulate physical activity. A walking distance of 1000-1600 metres to school has been found to contribute to overall levels of physical activity for boys [@faulknerSchoolTravelChildren2013]. The daily routine of travelling to school can be a good opportunity for children to regularly build physical activity into their schedule [@mitraIndependentMobilityMode2013]. Research has also shown that using active modes to school contributes to improve cardiovascular fitness [@borrestadExperiencesRandomisedControlled2012].

More recently, the literature has explored the link between transport and children's wellbeing [@waygoodTransportChildrenWellbeing2020], with relevant applications to the study of travel satisfaction [@vandenbergFactorsAffectingParental2020; @westmanChildrenTravelSchool2017]. Being driven to school reduces community interactions for children, which may negatively impact their social wellbeing [@waygoodChildrenTravelIncidental2015]. In contrast, walking and cycling give children opportunities to socialize with friends or siblings [@michailChildrenExperiencesTheir2021]. This is something that children seem to value [@zwertsHowChildrenView2010] and that tracks with findings among university students [@paezEnjoyment2010]. More generally, social connections through travel appear to be important for wellbeing. Furthermore, AST can provide opportunities for children to engage with natural environments [@fuscoUnderstandingChildrenPerceptions2012; @romeroChildrenExperiencesEnjoyment2015].

@@

Factors that influence AST have been presented and organized using a socio-ecological model (SEM) [@mitraIndependentMobilityMode2013] or systems model [@badlandDevelopmentSystemsModel2016] whereby children's travel behaviour is understood within the context of household, social, neighbourhood, and policy environments. The SEM comes from the field of public health and is a useful framework for understanding complex health behaviours, such as walking and cycling habits, This model helps to identify multiple determinants that need to be addressed by interventions to facilitate behaviour change. The consensus in the literature is that interventions should target multiple levels in order to increase levels of healthy and physically active modes of travel to school [@mitraIndependentMobilityMode2013]. Rather than describing an extensive list of factors that influence AST and mode choice in Canada [e.g., @mammenUnderstandingDriveEscort2012; @mitraIndependentMobilityMode2013; @rothmanDeclineActiveSchool2018; @wilsonUnderstandingChildParent2018], we discuss a few potentially modifiable determinants that may be targeted for change through STP.

At the individual level, age is often positively associated with AST [@mammenUnderstandingDriveEscort2012; @starkExploringIndependentActive2018; @wilsonUnderstandingChildParent2018]. There is evidence that gender is a determinant of AST, and that boys are more likely to travel using active modes than girls, although this is not a strong or consistent finding [@rothmanDeclineActiveSchool2018; @schoeppeAssociationsChildrenActive2015]. Children's mode choice to school is strongly influenced by their parents' travel behaviours and the complexity of their household's travel needs [@buliungLivingJourneySchool2021], as well as the frequency of parental support [@mahDoesParentalSupport2017a]. This indicates that shifting parental perceptions and habits is important. Convenience and inclement weather have been cited by parents as barriers to AST [@buliungSchoolTravelPlanning2011]. Parental perceptions of the built or school environment [@demeesterParentalPerceivedNeighborhood2014; @panterAttitudesSocialSupport2010] and their children's skills [@mammenUnderstandingDriveEscort2012; @faulknerWhatQuickestEasiest2010] also influence mode choice to school.

Distance between home and school is most strongly associated with AST [@ikedaAssociationsChildrenActive2018; @mammenUnderstandingDriveEscort2012; @pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018] with less AST reported among children who have to travel farther to school - low density suburban development tends to increase the need for movement, and the reliance on motorized modes [@farberRunning2011]. Many studies have also found that the quality of the built environment along the route to school and around the school site [@ikedaAssociationsChildrenActive2018; @rothmanActiveSchoolTransportation2021] and provision of active travel infrastructure [@chenPromotingActiveStudent2018; @pontEnvironmentalCorrelatesChildren2009] facilitate AST. Canadian youth report that they feel most safe bicycling on streets in their neighbourhood or that have low volumes of traffic [@TACbikeinfra2020]. Finally, concerns about traffic and strangers have been reported by parents who drive their children to school [@mammenUnderstandingDriveEscort2012], which highlights that the volume and speed of cars can be a concern or deterrent for AST.

@@

How AST is framed seems particularly important to shift parental attitudes and perceptions given their reported influence on children's travel mode to school. STP activities heavily focus on education or encouragement [@buliungSchoolTravelPlanning2011; @mammenActiveSchoolTravel2014; @buttazzoniSupportingActiveSchool2018], but parents may not always be receptive to the goals of STP and may be resistant to behaviour change [@buttazzoniSupportingActiveSchool2018]. Parents have been found to express different understandings, language, and perceptions than planners of how the built environment can influence school travel [@buliungLivingJourneySchool2021]. This is also true when it comes to other factors like convenience of different modes to school [@langUnderstandingModalChoice2011]. STP stakeholders must pay special attention to parents' understanding of the decline of AST as a problem, which may affect their receptivity to proposed solutions.

The “central organizing idea or story line" of AST, to apply the definition of framing from Gamson and Modigliani [-@gamsonChangingCulture1987], could also affect broader support in the community. Municipal representatives are perceived to be instrumental but the involvement of other stakeholder groups (e.g., busing consortia representatives and local residents) can be lacking [@buttazzoniSupportingActiveSchool2018]. In Ontario, it is reasonable to say that AST has become a policy issue on the education and public health agendas, supported, albeit in a modest way, by financial contributions from the provincial government^[To put the CAD 2 million invested by the province in the Ontario Active School Travel program, a city such as Hamilton budgeted in excess of CAD 91 million on road maintenance and construction in 2016 alone; see: https://www.hamilton.ca/budget-finance/city-budgets/2016-tax-and-rate-budgets]. Still, the support from a range of municipal representatives [see @buttazzoniSupportingActiveSchool2018; @mammenPuttingSchoolTravel2015] demonstrates that the AST issue is on the political agenda. However, it is unknown to what degree the general public (i.e., local residents) has been exposed to messaging about AST.

@@

kable(data_tbl,
      "latex",
      booktabs = TRUE,
      caption = "\\label{tab:search-results}Search results from the main STP stakeholder groups.") %>%
  kable_paper(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(3, border_left = T) %>%
  kable_styling(latex_options = c("striped"))
```

@@

We conducted a search on Web of Science for scholarly papers on the topic of travel to school. We limited our search to the fields of public health, transportation, planning, urban studies, and geography. This initial corpus was curated by the authors to ensure that all documents were relevant; for example, some hits in our search were not full papers but conference reports, or they dealt with travel to school only peripherally. After this step, we assembled a collection of 227 journal articles for analysis.

@@

A multi-step process was conducted to ensure that the analysis captured as much text as possible from both the policy documents (n = 64) and academic papers (n = 227). To begin, the webpages, which were manually downloaded in portable document format (PDF), were trimmed so that pages that only consisted of tables, figures, or references were removed. Many academic papers were in a two-column format, which is not ideal for conversion to `txt`. Two-column PDF documents were processed to extract the text in the correct sequence. Four academic papers did not join sufficiently and were taken out of the corpus due to the substantial time required to manually correct their inconsistencies.

Next, we converted the trimmed PDF documents into `txt` files so that they could be imported for analysis in `R`. The next phase was to manually remove any remaining tables, figures, references, headers/footings, and captions that could not be trimmed. Manual corrections were also required for certain pages in academic papers that remained in two-column format after the conversion process. This typically occurred on pages that had a table or figure that disrupted the text. Finally, we reviewed all of the documents to remove hyphenation by line breaks and to keep hyphenated words together on the same line. Any ligatures (e.g., combinations of characters or letters that were not properly detected during the conversion process) were fixed.

We also manually removed any extraneous material in the academic papers that did not pertain to AST specifically. This included footnotes, references, acknowledgments, and conflict of interest statements in the academic papers. We removed all phone numbers, inserted links to other webpages, personal names, and content not specific to AST from the policy documents that were retrieved from the websites of school boards, municipalities, and transportation consortia.

@@

We use topic modelling in `R` to conduct the framing analysis. Topic modelling is a machine learning technique used to analyze text to identify the language and concepts being communicated. This method is practical for researchers working with large amounts of text because it replaces the manual coding of topics that would normally take place to analyze or summarize textual data [@jacobiQuantitativeAnalysisLarge2016]. In the data pre-processing phase, we tokenize the text in the documents and create a document-term matrix so that it is in the correct format for analysis. We primarily use the following packages: `tidytext` [@R-tidytext], `topicmodels` [@R-topicmodels], `word2vec` [@R-word2vec], and `wordcloud` [@R-wordcloud] to examine text in the documents that were sourced for this project. These packages have functions for determining the frequency of specific words in each document or relationships (e.g., pairs of adjacent terms called bigrams) and correlations between words. Topic modelling is a popular method for analyzing text from social media platforms [@albalawiUsingTopicModeling2020] and news articles [@jacobiQuantitativeAnalysisLarge2016]. We estimate latent Dirichlet allocation (LDA) models to classify both the STP and academic documents according to the topics that are contained within them. This method "treats each document as a mixture of topics, and each topic as a mixture of words" [@silgeTextMining2021]. The model's output is "a set of topics consisting of clusters of words that co-occur in these documents according to certain patterns" [@jacobiQuantitativeAnalysisLarge2016]. Researchers must then interpret the identified topics, as done after other methods of manual coding. We also compare the topics between the policy documents and the academic papers, and use our interpretation to answer our research question. We describe in more detail the functions and code in `R`  that we used.

@@

This paper is an example of open and reproducible research that uses only open software. All data are organized in the form of an open data product [@arribasOpen2021]. Following best practices in spatial data science [@brunsdon2020opening], the code and data needed to reproduce our research or conduct a similar analysis for other regions are available for download.

@@

We analyzed word and document frequency for each corpus of text. Table \ref{tab:word-table} shows the most frequent terms found in the municipal, transportation consortia, school board, and academic documents. Policy documents and academic papers reference _active_, _travel_, _walking_, _biking_ or _cycling_, and _students_ more than other terms. Each corpus also has _safety_ and _traffic_ as common words which suggests that such concerns are a common reason given for parents driving their children to school. The word _physical_ is present in each corpus, but it's not clear what this refers to (e.g, _physical activity_, _physical health_, or the _physical environment_). Furthermore, documents from STP stakeholder groups discuss _resources_, _information_, and _services_ about school travel. Unlike the academic papers, policy documents include the words _route_ or _routes_ as frequent terms. This could indicate the presence of "safe routes to school" in the policy corpus, as well as the role of STP stakeholder groups in identifying safe routes to school to share with parents or families. In the section below, the context in which these terms are used is explored further.

The academic corpus differs from the policy documents in that _parents_ and _distance_ are the second and third most common terms. In addition, _time_, _factors_, _environment_, and _age_ are also identified in the academic papers. The prevalence of these terms is consistent with an academic focus on exploring the variables that influence mode choice. These words are absent from the list of common words in policy documents. Table \ref{tab:word-table} indicates that the academic corpus discusses a broader range of determinants of AST than the policy documents. The number of references for each term in the academic papers is also substantially higher due to the inclusion of more documents.

@@

Examination of document frequency reveals terms that are not present in all policy documents. This suggests that although documents pertain to the subject of school travel, not all stakeholders across Ontario are disseminating information about AST. For example, a document may discuss active travel but not to school. Some documents present school travel options by bus but not by active modes. We manually searched the policy corpus and found that 48% of documents mention AST and 16% mention STP. This confirms that many municipalities, school boards, and transportation consortia do not promote AST through their webpages or indicate their involvement in the STP process. Instead, inclement weather and its impacts on busing is a common topic addressed in school board and transportation consortia documents.

@@

Bigrams refer to a pair of consecutive words. Figures \ref{fig:city-visual}, \ref{fig:consortia-visual}, and \ref{fig:school-visual} show the bigrams that occur more than 5 times for each set of policy documents. These figures help to make further sense of the word frequencies reported above, and highlight the main ideas that are presented to the public in each of the policy corpora. The directional arrows indicate the arrangement of the words (e.g., active travel and not travel active) and the colour gradient of the arrows corresponds to the most frequently mentioned pairs (e.g., bigrams with darker arrows are found more often).

Municipalities primarily discuss _physical activity_ (n = 53) and _public health_ (n = 19) in the context of AST. In addition, _travel planning_ (n = 19), _bike lanes_ (n = 16), and _safe routes_ (n = 14) are also identified, conceivably as either proposed solutions or built environment factors that support AST. Key issues related to transport such as _traffic safety_ (n = 10), _air quality_ (n = 9), and _greenhouse gases_ (n = 9) are conveyed to the public through these policy documents. It is not surprising to find this focus given that municipalities in Ontario are concerned about climate change and have increasingly looked to active travel to offset transport-related emissions in urban areas.

Similar bigrams are found in school board documents: _travel planning_ (n = 33), _safe routes_ (n = 15), _physical activity_ (n = 10), and _public health_ (n = 10) are among the most common. Both municipalities and school boards in Ontario seem to emphasize what can be or has been done to improve AST (i.e., policy or planning changes), while outlining some of the benefits of AST at the individual- or community-level to potentially encourage behaviour change (i.e., physical activity for children or improved air quality). Unlike other STP stakeholders, school boards also consider _inclement weather_ (n = 24) and _bus cancellations_ (n = 13). Most likely this is because school boards are mandated to provide transportation to school in Ontario (typically by bus) and this information is presented alongside AST options. Finally, transportation consortia documents highlight topics such as _physical activity_ (n = 10), _pedestrian safety_ (n = 8), _crossing guards_ (n = 6), _travel planning_ (n = 6), and _walk zones_ (n = 6). Biking and cycling are notably absent from transportation consortia documents. Overall, the policy documents appear to convey an emphasis on the built environment, rather than household decision-making.

@@

We interpreted the most common bigrams from the policy+practice corpus (see Figure \ref{fig:policy-visual}), which includes all documents from municipalities, transportation consortia, and school boards, as the main ideas that STP stakeholder groups focus on and communicate to the public about AST. We then examined the context of these key ideas by extracting words-in-context from the corpora. Table \ref{tab:policy-concordance} presents some examples from select policy documents that illustrate how the most common bigrams are communicated to the public.
